---
title: 'Stack Overflow Developers Survey: Data for Greek Developers'
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

The two datasets that I use originate from the "Stack Overflow" website.
These two datasets are the result of the yearly survey conducted by the
company, regarding developers all around the world.
(<https://insights.stackoverflow.com/survey/2021>). I used the dataset
from 2021. More information regarding the datasets can be found online.

Insert the library "tidyverse" and our datasets.

```{r}
library(tidyverse) 
survey <- read.csv("survey_results_public.csv")
schema <- read.csv("survey_results_schema.csv")
```

View the datasets, check dimensions and column names.

```{r}
dim(survey)
view(survey)
head(survey)
colnames(survey)
view(schema)
```

In this small case study, I want to take a look at data relevant to
greek developers. I will filter out only the developers that live in
Greece.

```{r}
greece <- survey%>%
  filter(Country == "Greece")
head(greece)
```

In contrast with other countries, the amount of respondents from Greece
is smaller. We can check how many observations there are with the
n_distinct() function.

```{r}
n_distinct(greece$ResponseId)
```

The new dataset has 603 responses. Let's check for duplicates before
proceeding.

```{r}
isTRUE(duplicated(greece))
```

We see that we get one False value, meaning that we do not have
duplicate entries. We can check for the same question without adding the
"is.TRUE()" function. In that case, R will give back a number of Falses
and/or Trues.The new data set that I created includes 603 developers
that live in Greece. However, I want to extract answers about those that
are professionals. I will filter out only those devs that code
professionally and are employed full time.

```{r}
pros <- filter(greece, MainBranch =="I am a developer by profession" & Employment =="Employed full-time")
head(pros)
n_distinct(pros)
view(pros)
```

The new data set is ready. For start, I want to check the median salary
of developers living in Greece

```{r}
median(pros$ConvertedCompYearly)
```

It seems that we have some respondents that did not include their
salary. We have to clean the data set and keep only those responses with
actual information. In this case I will remove the NA values. However,
another good practice that suits proper analysis would be the
replacement of NAs with the MEAN or the MEDIAN of the rest of the
observations. Also, let's remove some columns that are not needed.

```{r}
pros_omit <- pros%>%
  drop_na(ConvertedCompYearly)
pros_omit <- pros_omit%>%
  select(-c(US_State, UK_Country))
```

Let's check the median salary of the developers living in Greece.

```{r}
median(pros_omit$ConvertedCompYearly)
```

The median salary of developers living in Greece is 25944 Euros. If we
take a look at our data set, we will notice that there is a column which
mentions the programming languages these devs have worked with. For the
sake of this small case study, let's create a column with the number of
languages each dev has worked with and know, and let's see if there is a
correlation with their yearly salary.

```{r}
pros_omit$lang_work <- gsub(";", " ", pros_omit$LanguageHaveWorkedWith) ## create a new column and replace ";" character with a space " "

langs <- pros_omit$lang_work ## create a new vector with the separated values
langs2 <- unlist(sapply(langs, strsplit, split = "\\  ", USE.NAMES = FALSE)) ##unlist the vector langs2 
pros_omit$langs_number <- (str_count(langs2, " ")) ##assign the number of strings as a number in our dataframe

pros_omit$langs_number <- pros_omit$langs_number + 1 ## add 1 in the column for correcting the missing language
                           
view(pros_omit)
```

Let's see if there is a correlation between the number of languages used
by devs and their yearly income

```{r}
ggplot(pros_omit, aes(x = langs_number, y = ConvertedCompYearly)) + 
  geom_point()+
  geom_smooth(method = lm)+
  ggtitle("Corellation between number of languages known and annual income")+
  labs(x = "Languages known", y = "Annual income")
```

This data set includes some extreme values, such as 1 million or 2
million Euros. For reasons of simplicity let's narrow down these values
by setting a filter to yearly income and re-plot our results.

```{r}
pros_inc <- subset(pros_omit, pros_omit$ConvertedCompYearly < 80000)
view(pros_inc)

```

Create the scatter plot again

```{r}
ggplot(pros_inc, aes(x = langs_number, y = ConvertedCompYearly)) +
  geom_point(col = "red", pch = 15)+
  geom_smooth(method = lm)+
  labs(x = "Languages worked with", y = "Yearly Income")
  
```

By plotting the data, it is obvious that there is not a direct
correlation between the number of languages known and previously used by
developers and their annual income. Now, we will explore the potential
correlation between income and years coding and age. First we have to
clean those columns in order to explore our data. Then we will create a
new plot

```{r}
pros_inc%>%
  na.omit(pros_inc$YearsCodePro)
isTRUE(is.na(pros_inc$YearsCodePro))
pros_inc$YearsCodePro[pros_inc$YearsCodePro == "Less than 1 year"] <- 1 ## replace "Less than 1 year" string with 1
view(pros_inc)
pros_inc$YearsCodePro<- as.numeric(pros_inc$YearsCodePro)
view(pros_inc)
ggplot(pros_inc, aes(x = YearsCodePro, y = ConvertedCompYearly))+
  geom_point(col = "red", pch = 15)+
  geom_smooth(method = lm)+
  ggtitle("Correlation between years of professional coding and annual income")+
  labs(x = "Years of professional coding", y = "Annual income")

```

In this particular case, a small trend seems to exist between the two
variables, as in some cases, more years of coding experience lead to a
better yearly salary. However, this result definitely does not provide a
secure answer and more sophisticated methods should be applied, which
lie beyond this small case study.

Now, let's create some visualizations in order to explore our data

Density of developer's age groups based on their yearly income:

```{r}
ggplot(pros_inc, aes(x = ConvertedCompYearly, fill = Age))+
  geom_density(alpha= .7)+
  ggtitle("Density of programmers age groups based on their income")+
  labs(x = "Annual income", y = "Density")
```

It is obvious that most developers between 18-24 years old earn less
than 20k euros per year, with only a small number of them earning more
than this amount. The graph gives some excellent result about Greek
professional developers which need to be explored in more depth.

In the same way, we can further explore our data and answer some simple
(at the beginning) questions regarding the situation with Greek
professional developers. For example, we can check with some simple
plots the differences between gender, education or other similar issues.

```{}
```
